### 2021-09-14 paper/outline.tex
added outline.tex

\usepackage{outlines}
\renewcommand{\outlineii}{cenumerate}
\begin{outline}

\end{outline}