### 2021-11-27 paper/abstract.tex
abstract

\documentclass[11pt]{article}
\usepackage{xcolor}
\usepackage{times}
\usepackage[margin=1in]{geometry}

\begin{document}

\noindent \textbf{CAKE: Computational and Automated Kantian Ethics}

\medskip

\noindent \textbf{Abstract}: As artifically intelligent agents become increasingly powerful, the need for philosophically
sophisticated automated ethical reasoning becomes more pressing. Representing a robust ethical theory 
to a computer poses technical and ethical challenges. I argue that, out of the three major ethical traditions, Kantian ethics is
most natural to formalize because it presents inviolable, context-agnostic, formal rules. In this paper, 
I present CAKE, an implementation of automated Kantian ethics that is faithful to the rich Kantian 
philosophical tradition. I formalize Kant's categorical imperative in deontic logic, implement this 
formalization in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics. Not only is CAKE an early step 
towards philosophically rich ethical AI agents, it also establishes that computational tools can help 
philosophers reach ethical insights.

\smallskip
\noindent \emph{Areas: algorithm development, applications, philosophy}
\end{document}