### 2021-12-13 paper/abstract.tex
paper outline

\documentclass[11pt]{article}
\usepackage{xcolor}
\usepackage{times}
\usepackage[margin=1in]{geometry}

\begin{document}

\noindent \textbf{A Faithful Implementation of Automated Kantian Ethics}

\medskip

\noindent \textbf{Abstract}: Warnings from regulators, philosophers, and computer scientists about the 
dangers of unethical artificial intelligence have spurred interest in the development 
of machines that can perform ethical reasoning. Despite the fact that ethical philosophers are devoted to 
developing robust ethical theories, previous work in automated ethics rarely engages with existing 
philosophical literature. Faithfully translating complex, philosophically sophisticated ethical theories 
expressed in natural language to the rigid syntax of a computer program poses technical and philosophical 
challenges. In this paper, I present an implementation of automated Kantian 
ethics that is faithful to the Kantian philosophical tradition. Of the three major ethical 
traditions, Kant's categorical imperative is the most natural to formalize because it is an inviolable, 
context-agnostic, formal rule.  I formalize Kant's categorical imperative 
in Carmo and Jones's dyadic deontic logic, implement this formalization 
in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics, as established in the literature. 
My system is not only an early step towards philosophically mature ethical AI agents, but it can also help
philosophers reach new ethical insights, thus paving the way for the use of computational tools to 
address philosophical problems.

\smallskip
\noindent \emph{Areas: algorithm development, applications, philosophy, ethics and automation}
\end{document}