### 2022-02-04 paper/outline_notes/notes_01_13.tex

Rawls Lectures on the History of Moral Philosophy p. 166
"It is a serious misconception to think of the CI-procedure as an algorithm intended to yield, more 
or less mechanically, a correct judgment. There is no such algorithm, and Kant knows this. It is 
equally a misconception to think of this procedure as a set of debating rules that can trap liars
and cheats, scoundrels and cynics, into exposing their hand. There are no
such rules."
- why not? rawls doesn't really explain himself

O'Neill Universal Laws and Ends in Themselves p. 343
"The three formulations are all offered as tests that agents can apply to
 proposals for action. The Categorical Imperative is nowhere proposed as a
 principle that will by itself generate or entail a universal moral code. It is not
 a moral algorithm (unlike the Principle of Utility) but (supposedly) a
 criterion of moral action for agents who act freely, so may start with
 various possible proposals for action. The common assumption of the three
 principles is that there is some way by which agents can filter these initial
 proposals to check whether they are morally acceptable. Each formulation
 of the Categorical Imperative is offered as an answer to the agent's question
 "What ought I do?",3 on the assumption that agents will have certain ten
 tative plans, proposals and policies which they can consider, revise or re
 ject?or endorse and pursue."
her arguments:
-can only test given maxims, can't magically produce moral action 
-assumes that there is some input and answers the question "What should I do?" among given options
  - that is exactly what my system does 
  - the model is that, like a human, an AI navigating the world has options and has to decide among them

p.345
"Fifth, since the implementation of maxims will differ according to cir
 cumstances, a test on maxims is not and cannot be enough to determine the
 Tightness or wrongness of particular acts (their "legality"); it can only
 reveal the moral quality or worth of maxims (so is in Kant's terms a test of
 "morality"). Kant defines duty not (as would be common today) as out
 ward performance of a certain sort, but as action that embodies a good will
 i.e., action on a maxim of a certain sort. (397) However, although moral
 worth is more fundamental than Tightness in Kant's theory, Tightness and
 wrongness are more easily ascertainable. This is simply a corollary of the
 opacity of our self knowledge. If we are unsure what the maxim of a given
 act is, we cannot be sure whether it is morally worthy. Despite their best ef
 forts at principled and self-conscious action, agents are prey to self
 deception and selective perception. This is not rare or exotic but com
 monplace?we are repeatedly tempted to ascribe maxims that place acts and
 agents in a more flattering or a more lurid light. By contrast, it would be
 relatively easy // we had a test to identify morally worthy maxims to deter
 mine whether an agent who acted on such a maxim would have acted in a
 specific way in those circumstances. In his most pessimistic moments Kant
 doubts whether we can ever know that a morally worthy action has been
 done, which would mean that we can never judge the morality of acts. This
 pessimism need not stop us from judging whether acts that conform to such
 maxims have been done?provided that we have a criterion for identifying
 morally worthy maxims.5 The various formulations of the Categorical Im
 perative are supposed to provide this criterion"
-only evaluates whether or not maxims are morally worthy, but doesn't evaluate action. to evaluate 
action, you have to decide what the maxim of action is and evaluate it
  - in the model above, the AI has options and has to decide which to take, like a human. Section ??
concedes that determining the maxim to pass as input to my system is a challenge and possible limitation
of my system, and this is part of the challenge of formulating a maxim, which this criqitue also 
says is a challenge.
  - Nonetheless, there is some value to providing a criterion for identifying morally worthy maxims. 
Perhaps we (and machines) are not always mistaken about the maxim of action. Perhaps the maxims are chosen
from some pre-populated, manually entered database. Identifying morally relevant maxims is one step 
towards automated ethics but not the whole puzzle and that's ok. 
p.349-350 describes what the role of the FUL is 
- most practical reasoning doesn't involve literally using the CI to evaluate maxims - we have some 
prior knowledge from moral advice and guidance about which maxims are good or bad
- FUL merely tells you how to distinguish maxims of duty from maxims that are not of duty, but practical
reason is the process of determining which maxim a particular action in a particular situation conforms with
  - this provides a clue as to how you can implement an input parser - maybe we have some prior database of 
actions or action-maxim mappings and we extrapolate from there 
  - either way, we agree. my project tells you how to distinguish maxims of duty, and then actually 
using that to guide action requires an additional step
  - moreover, human beings need an "almanac" or prior set of maxims that are of duty or not because we
have limited reasoning capacity but computers can totally just rederive which maxims are good or bad
constantly. that way computers can use both testimony and proof!
- tradeoff: computation vs common sense so partnership is the best approach

seems crazy but need an argument given computes can simulate human brain

herman (available for pickup) TBD

Tafani
p3 "In Kant's lexicon, it is a matter of constructing machines whose
characteristic is not goodness, but holiness, that is, knowledge of moral
norms and the infallible and automatic adaptation of their actions to them"

Kant is easy bc:
p. 4 -> only requires rationality to behave morally (cognitivism)
no moral dilemmas bc no conflicting obligations
FUL seems computational

three challenges:
1. difficult/impossible for computer to check for contradiction under FUL
2. "the difficulty of selecting, from a given maxim, the relevant elements for its
universalization" (circumstance/formulation of the maxim)
3. how do you actually understand the universalized world -> need empirical and psychological knowledge
I solve the first and (kind of) second challenges
"universal diffusion of lying" is an empirical or anthropological truth
Herman -> moral rules of relevance -> human agent knows which features of the maxim are morally relevant

p. 8 -> Kant says that people know their duty and we need philosophy to remind them and mitigate bad-faith efforts
"For Kant, the identification of the criterion of moral judgement is therefore
not a procedure by which someone who is ignorant of the moral law can
find out what his duty is (Singer, 1973, p. 298), or by which a computer
system can be programmed to morality; it constitutes the analysis of
common moral judgement, i.e. the reconstruction of what every man
automatically does when he finds himself morally evaluating a maxim. The
Kantian test is therefore as useless for machines as it is for anyone who does
not already know what to do."

his examples are cases where there is a temptation to make an exception just for me 
it's a reaffirmation in a moment of weakness, not a way to derive duty from scratch
"However useful it may be from a psychological
point of view, in moments of weakness, such a reaffirmation of the
universal validity of the moral law does not, by itself, lead to the definition
of what, concretely, must be done, since it is compatible - as a mere
requirement of universality - with countless alternative definitions of the
content of the moral law"

p. 8 "It is true - as O'Neill observes - that a check on the contradictory nature of
the universalised maxim requires a common sense consideration of the
empirical consequences, which a machine is therefore unable to perform,
and it is true that even a common sense test can give false positive results. It
is true that even a test based on common sense can give false positive
results. But this is not due to a weakness in the formulation of the test,
which would therefore need to be fine-tuned, but rather to a distortion of the
function attributed by Kant to the test by those who attempt to use it to
regulate the behaviour of humans or machines from scratch. "
- O'Neill says that you need some common sense or practical reason or judgement to perform the UT
- this is true and also why it gives false-positives
- BUT the false-positives are a result of using the test WRONG, in a bad-faith manner
- you CAN arbitrary specify maxims.but you're not supposed to
- you're not supposed to try to circumvent the test - the point is not to put in arbitrary maxims
- instead, the point is to put in maxims of which you have some notion of rightness or wrongness
and the test will confirm it and remind you what your duty is

outline:
1. philosophers like to assert that ethics is not the right kind of thing to be automated (e.g. Rawls). 
      But this has to be justified - after all, computers can mimic human brain activity entirely, so cognitivists
      have to argue why ethics is out of reach of computation but addition, emotion, thought, and language aren't

2. one reason, from O'Niell is that the FUL can only test given maxims but can't actually prescribe
moral action because moral action also requires a specific attitude towards the moral law
      That's ok! Current computers aren't rational beings so they can't actually behave morally, but
      they can act on the maxim that has moral worth in a particular situation. We can't judge a computer
      as "acting from duty" but we might still think that a computer acting on morally worthy maxims 
      is a better thing for society than computers not acting on morally worthy maxims.

      Also, this doesn'tindict CE.

3. Moreover, even evaluating a particular maxim requires that the maxim be given as input. Specifically,
human beings navigating the world are faced with choices, and if they want to use the FUL, they have to
extract their potential maxims from their choices. O'Neill says 
that even Kant identifies this as challenging for human beings, and Tafani also identifies this as a 
challenge for Kantian machine ethics.
      This is a challenge I also identify! Formulating the input maxim for my system. See the discussion 
      in Section Upshot about this. Roughly, I argue that either a human being can take over or we can 
      develop heuristics. Moreover, there's still some value to having a machine that can apply the CI
      test to given maxims and future work can work on the input parser. Also doesn't indict CE.

4. Finally, Tafani and O'Niell also argue that the FUL is not SUPPOSED to generate ethics from scratch.
O'Niell notes that human beings come equipped with an almanac of right and wrong maxims and that
the fUL tests these. Human beings don't actually repeatedly apply the UT.
      Cool! Computers don't have human limitations, so they can actually repeatedly apply the CI test, 
      essentially checking their work and rederiving the moral worth of a maxim every time they consider it.
      There's a tradeoff here: humans have common sense but computational limits, so they use the common 
      sense to avoid repeating calculations. computers have no common sense but (Compared to humans)
      no computational limits, so they can perform the calculation over and over again instead of
      "remembering" it. almanac == cultural memory of rightness and wrongness

5. Tafani takes it one step further by arguing that we actually know which maxims are right
and which ones are wrong and the UT is just supposed to strengthen our resolve to behave morally whenever
we are tempted to make an exception of ourselves. Indeed, bad-faith attempts to pass the UT work - see 
the example of the tailoring objection given by O'Niell. 
    This, combined with the almanac, actually give us some insight into how we can solve the challenge
    of making an input parser. Specifically, we can learn some set of maxims and maybe assign each a 
    "rightness score" indicating how confident we are that it is right or wrong and THEN use the automated
    FUL to test these maxims. You can imagine a deep-learning type system that, based on some database
    mapping actions to maxims, can extract a maxim from a situation as well as its rightness score
    and then pass these as input to the FUL test. The output of the FUL test can in turn reinforce or
    rething the rightness score.