### 2022-03-21 paper/output/document/thesis_0_acknowledgements.tex

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}acknowledgements}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
Every step of this process has been made easier by my mentors' unwavering support. Thank you to Professor 
Nada Amin for taking me seriously when this project was just a vague intuition, for always asking the 
right questions at the right time, and for being willing 
to stumble through the dark with me. I am endlessly grateful for our Friday night email threads.

Thank you to Dr. William Cochran for offering much more of time, energy, and support than I ever 
imagined possible. This project was less lonely because of your
willingness to venture down rabbit holes with me. I hope that someday I can be as cool and 
collected as you (maybe I'll have to read more Aristotle first).

Thank you to Professor James Waldo for always giving me the advice I need to hear. Since freshman year, 
I have always walked out of your office with a clear head and renewed resolve. Thank you for giving me the 
knowledge that, even when everything feels hopeless, there is one person at this university who believes in me.
Your mentorship has been the single most transformative part of my college experience, and I will always
be grateful to you.

I fell in love with philosophy while reading Christine Korsgaard's \emph{Sources of Normativity}, and 
this thesis is largely inspired by her work. She once said that 
we make the world more sensible by being good to each other; her ideas have certainly made my world make more
sense.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_0_acknowledgements.thy%:%
%:%19=6%:%
%:%20=7%:%
%:%21=8%:%
%:%22=9%:%
%:%23=10%:%
%:%24=11%:%
%:%25=12%:%
%:%26=13%:%
%:%27=14%:%
%:%28=15%:%
%:%29=16%:%
%:%30=17%:%
%:%31=18%:%
%:%32=19%:%
%:%33=20%:%
%:%34=21%:%
%:%35=22%:%
%:%36=23%:%
%:%37=24%:%
%:%38=25%:%